\section{Introduction}
\label{sec:introduction}

Writing unit tests manually is a time-consuming and error-prone process, but complete test coverage remains essential for detecting logical errors before program release. For functions with a cycle complexity greater than 5, the range of forks becomes so broad that developers must meticulously list boundary states, boundary conditions, and exception paths---a task that is geometrically time-consuming compared to the amount of code being tested. The need for scalable automated test generation is therefore well-recognised in both industry and research.

Recent studies have examined large language models (LLMs) as a promising alternative to traditional search tools such as EvoSuite and Randoop. Sch{\"a}fer et al.~\cite{schafer2024testpilot} have shown that \texttt{gpt-3.5-turbo} achieves an average of 52.8\% branch coverage for JavaScript, TypeScript, and Python functions using the command line without training. Yuan et al.~\cite{yuan2024chattester} have also shown that reloading a large language model with compiler feedback within a single revision cycle increases branch coverage to over 82\% at the Java class level. Lops et al.~\cite{lops2025agonetest} reported up to 79.8\% branch coverage across three model configurations applied to a Java benchmark. Celik and Mahmoud~\cite{celik2026gem} found that LLM-generated tests outperform Randoop on mutation score when evaluated on isolated functions. These studies establish that LLMs can produce high-quality tests; however, direct comparisons between them are confounded by differences in generation protocol (single-shot versus iterative repair), level of analysis (method versus class), and the structural characteristics of the evaluated code.

No prior study has simultaneously (a)~filtered subject functions to a controlled cyclomatic complexity band, (b)~evaluated a single first-attempt response without iterative repair, and (c)~compared against a classical automated baseline within a pre-registered statistical design. As a result, it remains unclear how code complexity affects first-attempt LLM test quality, and whether the observed improvements over random tools hold when complexity is held constant.

In this paper we address this gap by evaluating \texttt{gpt-4o-mini-2024-07-18} on 50 standalone functions (25~Java, 25~Python) with cyclomatic complexity strictly between 5 and 15, using a single context-enriched prompt and no iterative repair. We measure branch coverage and mutation score under a pre-registered analysis plan, and we compare the Java results directly against Randoop~4.3.4 using a paired design.

To summarise, this paper contributes:\\
(1) The first pre-registered, complexity-controlled empirical evaluation of first-attempt zero-shot LLM test generation on a balanced Java/Python sample, measuring both branch coverage and mutation score.\\
(2) Experimental results showing that \texttt{gpt-4o-mini-2024-07-18} achieves median branch coverage of 95.5\% (Java) and 100\% (Python) on CC~5 to 15 functions, with both populations rejecting the 75\% null at $p < 0.001$.\\
(3) A paired comparison against Randoop on Java mutation score demonstrating a 62.1 percentage-point median advantage for GPT-generated tests (Cliff's $\delta = 0.79$, $p < 0.001$).\\
(4) A public replication package including the frozen manifest, all API request/response logs, and the pre-registered analysis script at \url{https://github.com/phamvohungnghi/SWT301-SE1944-Group6}.

The rest of this paper is structured as follows. Section~\ref{sec:related_work} reviews related work on LLM-based and search-based test generation. Section~\ref{sec:methodology} describes our experimental design, dataset, model configuration, and statistical analysis plan. Section~\ref{sec:results} reports the results for each research question. Section~\ref{sec:discussion} interprets the findings and compares them with prior work. Section~\ref{sec:threats} discusses threats to validity. Section~\ref{sec:conclusion} concludes and outlines future work.
