\section{Related Work}
\label{sec:related_work}

Automated unit test generation using Large Language Models (LLMs) has emerged as a promising alternative to traditional search-based and random testing techniques. This section reviews recent advances in this domain, structured into two key themes: LLM-based zero-shot unit test generation and feedback-driven iterative test refinement. We then position our current study within the identified gaps in existing literature.

\subsection{LLM-based Zero-shot Unit Test Generation}
\label{subsec:zero_shot_gen}

Initial investigations into the capability of LLMs for unit test generation focused primarily on their immediate, or \textit{zero-shot}, ability to produce tests given a production code unit without any iterative feedback or refinement. Guilherme and Vincenzi~\cite{guilherme2023sast} conducted an early study using \texttt{gpt-3.5-turbo} on 33 Java programs, reporting an average code coverage of 90.2\% and a mutation score of 70.5\%. Their work suggested that well-structured prompts could achieve a build success rate of 64.6\% with a \texttt{temperature} of 0.2, demonstrating early potential for LLM-driven test generation. Expanding on this, Schäfer et al.~\cite{schafer2024testpilot} presented TestPilot, an empirical evaluation involving \texttt{gpt-3.5-turbo}, \texttt{code-cushman-002}, and StarCoder on JavaScript/TypeScript and Python functions, primarily at the function/API level. They reported a median statement coverage of 70.2\% and a median branch coverage of 52.8\%. This study also highlighted challenges such as tests being dependent on specific contexts or external dependencies, and a notable presence of syntactic errors in the initial generation attempts.

Further evaluations of zero-shot capabilities were conducted by Celik and Mahmoud~\cite{celik2026gem} with their GEM framework, assessing models like Codex, Sonnet, and Qwen across Java, Python, and C++ at both function and class levels. They observed that Codex, when applied to the ULT(100) benchmark, achieved a branch coverage of 84.30\% and a mutation score of 72.71\%, with results varying significantly across different models and datasets. Similarly, Baskaran et al.~\cite{baskaran2025jucct} explored LLM-driven test case generation using OpenAI models orchestrated via the AutoGen multi-agent framework. Applied to a Python insurance management application, their approach achieved impressive branch coverage increases from 98\% to 99\% and mutation score improvements from 83.9\% to 95.8\%, though the scope was limited to a single application, potentially impacting external validity.

Despite these promising results, zero-shot generation faces critical limitations, particularly concerning the compilability and executability of generated tests on the first attempt. For instance, Lops et al.~\cite{lops2025agonetest} evaluated state-of-the-art LLMs, including \texttt{gpt-4o-mini} and \texttt{gemini-1.5-pro}, on the Java class-level CLASSES2TEST benchmark. While achieving maximum branch coverage of 79.8\% and mutation scores up to 89.2\%, the highest reported compilation rate was only 36.0\%. This highlights a significant gap between test generation quality and practical usability. Furthermore, many of these evaluations are conducted on general benchmarks without systematically controlling for specific structural attributes of the source code, such as cyclomatic complexity. This omission leaves a critical gap in understanding how intrinsic code complexity influences the quality and compilability of LLM-generated tests in a zero-shot setting.

\subsection{Feedback-Driven and Iterative Test Refinement}
\label{subsec:iterative_refinement}

To address the limitations of zero-shot generation, particularly compilation failures and suboptimal test quality, a distinct line of research has focused on integrating feedback loops and iterative refinement mechanisms into the LLM-driven test generation process. Yuan et al.~\cite{yuan2024chattester} proposed ChatTester, a framework that employs a multi-turn repair strategy to fix compilation and execution errors in Java class-level tests. This iterative approach demonstrated substantial improvements, increasing compilable tests by 34.3\% and achieving branch coverage exceeding 82\% alongside mutation scores above 65\%. This shows the effectiveness of post-generation repair in enhancing the practical usability of LLM-generated tests.

Similarly, Pan et al.~\cite{pan2024aster} introduced ASTER, a system that combines static analysis and context extraction techniques with advanced LLMs like GPT-4 to generate and remediate unit tests for Java and Python methods. On Python benchmarks, ASTER with GPT-4 achieved 78.0\% line coverage and 77.2\% branch coverage, along with positive feedback from a user study involving 161 developers. Their approach actively addresses runtime errors through remediation, contributing to higher success rates. Other frameworks directly utilize testing metrics as feedback. Straubinger et al.~\cite{straubinger2025mutation} developed an iterative scientific debugging approach for mutation testing, which successfully raised branch coverage from approximately 55\% to nearly 80\% and mutation scores from 60\% to 80\%. This method showcases how systematic debugging loops can significantly enhance test suite effectiveness. Additionally, Lu et al.~\cite{lu2026beyond} explored augmenting existing test suites using DeepSeek-V3 by specifically targeting survived mutants, leading to an average increase in mutation scores by 16.04\% for standalone methods and 8.11\% for non-standalone methods, highlighting the potential for LLMs to intelligently evolve test suites.

While feedback-driven and iterative techniques undeniably improve test compilability, coverage, and fault-detection capabilities, they often introduce significant operational overhead. Multi-turn interactions dramatically increase API token consumption, extend execution latency, and require robust integration with complex runtime verification and static analysis tools. Critically, by intertwining generation with extensive repair mechanisms, these studies do not isolate the base model's inherent zero-shot capability. Consequently, they often do not fully assess how intrinsic code complexity directly controls the quality and properties of the \textit{initial} LLM output, masking the underlying generation performance through subsequent refinement steps.

\subsection{Research Positioning}
\label{subsec:positioning}

Our study is positioned to address two primary gaps identified in the existing literature. Firstly, as highlighted in Section~\ref{subsec:zero_shot_gen}, a systematic evaluation of first-attempt (zero-shot) unit test generation performance of SOTA LLMs on a balanced dataset of Java and Python methods with explicit control over source code complexity (Cyclomatic Complexity between 5 and 15) is notably absent. Current benchmarks often include methods with arbitrary complexity, making it difficult to isolate the impact of complexity on LLM performance. Secondly, while several recent studies, such as Lu et al.~\cite{lu2026beyond} and Yuan et al.~\cite{yuan2024chattester}, have begun incorporating inferential statistics, many empirical evaluations still primarily rely on descriptive statistics. This leaves a methodological gap in rigorously establishing the statistical significance of observed differences in coverage and fault-detection capabilities between LLM-generated tests and traditional testing baselines. Our work directly contributes by filling these gaps: we systematically evaluate \texttt{gpt-4o-2024-08-06} in a zero-shot setting on a complexity-controlled dataset and apply rigorous non-parametric statistical hypothesis testing (specifically, the Wilcoxon signed-rank test and Mann-Whitney U test) to provide statistically sound conclusions regarding branch coverage and mutation score performance compared to traditional random testing baselines.
