\section{Methodology}
\label{sec:methodology}

This section describes our empirical study methodology, outlining the dataset
preparation, the large language model setup, the prompt engineering design, the
metrics measured, and the statistical analysis plan.

\subsection{Dataset and Subject Units}
\label{subsec:dataset}

We evaluate test generation on 50 standalone units: 25~Java methods and
25~Python functions.
All units were selected from public open-source repositories and from a shared
course project repository at
\url{https://github.com/phamvohungnghi/SWT301-SE1944-Group6}.

\textbf{Complexity filter.}
Cyclomatic complexity (CC) was measured using Lizard~1.23.0 on the source
revision identified by the commit SHA recorded in the dataset manifest.
Only units with $5 \le \text{CC} \le 15$ were eligible.
This band captures functions with genuine branching logic---too complex for a
trivial line-coverage sweep but small enough to be self-contained and runnable
without external services.

\textbf{Isolation requirement.}
Each unit must build and execute without network access, database connections,
wall-clock dependencies, or uncontrolled global state.
Two reviewers assessed isolation independently; any disagreement was resolved by
the project lead.

\textbf{Manifest and freeze.}
All eligible units were enumerated, assigned a \texttt{unit\_id}, and sorted by
the SHA-256 hash of the \texttt{unit\_id} string.
The first 25 Java and 25 Python entries were selected, yielding a deterministic,
cherry-pick-resistant sample.
The resulting manifest (\texttt{experiment/dataset-manifest.csv}) records, for
each unit: \texttt{unit\_id}, language, test framework, repository URL, commit
SHA, license, relative path, start/end lines, function name, Lizard CC score,
isolation status, SHA-256 source hash, and reviewer sign-offs.
The manifest was frozen and committed before any API call was made.

\textbf{Descriptive statistics.}
Cyclomatic complexity ranges from 5 to 15 for all 50 units by construction.
The Java units span 13 distinct CC values (5 to 15); the Python units span the
same range.
All functions are standalone: they accept and return primitive or standard-library
types without requiring test doubles for the production context.

\subsection{LLM-based Test Generation Setup}
\label{subsec:llm_setup}

For the automated test generation pipeline, we utilized the official OpenAI API
using the \texttt{gpt-4o-mini-2024-07-18} model snapshot.
To ensure the reproducibility of results and minimize bias, the parameters
generating the results are tightly controlled:
\begin{itemize}
    \item \textbf{Temperature:} Set to 0.0 to enforce deterministic code
    generation.
    \item \textbf{Top P:} The API requirements have been simplified to rely
    solely on temperature-based sampling.
    \item \textbf{Max Output Tokens:} The limit has been fixed at 2048 to avoid
    the possibility of the response being cut short.
    \item \textbf{System Prompt:} The system guided the model to operate like a
    software testing expert, adhering to the required framework, returning only
    executable source code, avoiding Markdown formatting blocks, and refraining
    from accessing external resources or modifying the production code.
\end{itemize}

At first (as documented in the revision log), we evaluated the process using
\texttt{gpt-4o} via the GitHub Models platform.
However, to address protocol discrepancies, collect accurate token consumption
data via the API \texttt{usage} field, and obtain standard OpenAI response
identifiers (\texttt{resp\_*}), we switched to the official OpenAI API and
reran all tests on 2026-07-16.

\subsection{Prompt Engineering and Context Injection}
\label{subsec:prompt_engineering}

A key challenge in first-attempt zero-shot unit test generation is ensuring the
generated code compiles and is syntactically correct the first time.
Our test evaluations show that without class structure and module context, LLMs
often produce invalid class definitions in Java or make incorrect import
assumptions in Python.

To overcome these issues, we created a prompt structure that includes context for
each language:
\begin{enumerate}
    \item \textbf{Language-Specific Instructions:}
    \begin{itemize}
        \item \textit{Java Instructions:} The model was told to call methods
        through the \texttt{JavaAlgorithms} class and forbidden from creating
        separate classes named after the unique unit identifiers
        (e.g., \texttt{JAVA\_001}).
        \item \textit{Python Instructions:} The model was told to import the
        target function using the syntax
        \texttt{from python\_functions import <function\_name>} and forbidden
        from copying the production code into the test file.
    \end{itemize}
    \item \textbf{Context Injection:} We included the entire target source file
    (or the first 200~lines if the file size exceeded 50~KB) as background
    context to provide the model with the exact imports, packages, and helper
    functions available.
    \item \textbf{Target Code:} The specific target method's source code was
    appended at the end of the prompt.
\end{enumerate}

\subsection{Execution Cost and Token Analysis}
\label{subsec:execution_cost}

To evaluate the performance and efficiency of our proposed framework, we
conducted extensive experiments using the official API endpoint.
A batch-processing run consisting of 60 successful requests was executed
(10~requests during the pilot phase and 50~requests during the full run),
consuming a total volume of 323,246 input tokens.
The average payload per request was approximately 5,387.4 input tokens,
indicating a highly dense contextual input/output interaction due to the
injection of source code context.
The total computational budget incurred during this empirical session was
\$0.0789~USD, demonstrating the high cost-effectiveness of our methodology.

Table~\ref{tab:token_cost} summarises the token consumption and estimated cost
using OpenAI's pricing structure for the \texttt{gpt-4o-mini-2024-07-18} model.

\begin{table}[htbp]
\centering
\caption{Token Consumption and Cost Details}
\label{tab:token_cost}
\resizebox{\columnwidth}{!}{%
\setlength{\tabcolsep}{3.5pt}
\begin{tabular}{lrrrr}
\toprule
\textbf{Phase} & \textbf{Input} & \textbf{Output} & \textbf{Total} & \textbf{Cost (USD)} \\
\midrule
Pilot (10 units) & 53,831 & 9,205 & 63,036 & \$0.0136 \\
Full Run (50 units) & 269,415 & 41,483 & 310,898 & \$0.0653 \\
\midrule
\textbf{Total (60 units)} & \textbf{323,246} & \textbf{50,688} & \textbf{373,934} & \textbf{\$0.0789} \\
\bottomrule
\end{tabular}%
}
\end{table}

\subsection{Metrics and Execution Tools}
\label{subsec:metrics}

We measure three primary outcomes for each generated test suite: executability,
branch coverage, and mutation score.

\textbf{Executability.}
A generated test suite is classified as \texttt{EXECUTABLE} if it compiles
successfully and all test methods complete without a JVM crash under JUnit~5.11.4
for Java, or under pytest~9.1.1 for Python.
Suites that fail to compile are classified as \texttt{COMPILE\_ERROR}; suites
that compile but abort at runtime are classified as \texttt{TEST\_RUNTIME\_ERROR}.
Every generated response, including failures, is retained in the denominator when
computing executability rates; no suite is silently dropped.

\textbf{Branch Coverage.}
For Java, branch coverage is measured at method granularity using JaCoCo~0.8.13
instrumented through the Maven Surefire plugin.
Only the target method's branches are counted; helper code in the same class is
excluded.
For Python, branch coverage is measured per function using coverage.py~7.14.2
with the \texttt{--branch} flag and a per-function \texttt{coverage run} scope.
Coverage is defined as
$\text{branch\_coverage} = \frac{\text{covered\_branches}}{\text{total\_branches}} \times 100\%$.
Non-executable suites receive \texttt{NA} in the primary complete-case analysis
and $0\%$ in the worst-case sensitivity analysis.

\textbf{Mutation Score.}
For Java, mutation score is measured using PIT~1.19.1 with its default mutator
set.
Timeout mutants are counted as survived in the primary analysis; invalid mutants
are excluded from the denominator.
Equivalent mutants are not removed by manual inspection after viewing results, to
avoid inflating scores post-hoc.
The formula is
$\text{mutation\_score} = \frac{\text{killed}}{\text{generated} - \text{invalid}} \times 100\%$.
For Python, pytest-mutagen~1.3 is used.
As a tool limitation, pytest-mutagen did not produce any valid mutants for any
of the 25~Python functions; consequently, Python mutation scores are \texttt{NA}
throughout and the Python mutation outcome is excluded from the confirmatory
analysis.

\textbf{Randoop Baseline.}
Randoop~4.3.4 is run for each of the 25~Java methods with
\texttt{randomseed=20260621} and a time budget of 60~seconds per unit, using
JDK~17 and the same production source revision as the GPT-generated tests.
Randoop-generated suites are passed through the same JaCoCo and PIT pipeline
without any manual repair.
Raw command logs, generated test files, and per-unit coverage/mutation reports
are archived in \texttt{evaluation/randoop/}.

\subsection{Statistical Analysis Plan}
\label{subsec:statistics}

All hypotheses are registered in \texttt{team-synthesis/proposal.md} before data
collection.
No threshold, test, or metric definition is altered after observing experimental
results.
The family-wise significance level is $\alpha = 0.05$.
We report raw $p$-values, Holm-adjusted $p$-values, effect estimates, and 95\%
confidence intervals for all confirmatory tests.

\textbf{RQ1 (Branch Coverage vs.\ 75\% Target).}
We apply a one-sided exact sign test (right-tailed) separately for Java and
Python, with $H_0$: median branch coverage $\leq 75\%$.
Observations exactly equal to 75\% are excluded as ties.
The two resulting $p$-values are adjusted for multiplicity using the
Holm--Bonferroni procedure.

\textbf{RQ2 (GPT-4o vs.\ Randoop, Java Only).}
We apply a one-sided paired Wilcoxon signed-rank test (right-tailed) on the
per-unit differences $d_i = \text{mutation\_GPT}_i - \text{mutation\_Randoop}_i$,
with $H_0$: $\text{median}(d) \leq 0$.
Zero differences are handled with the Wilcox zero-method.
As a sensitivity analysis, we apply a one-sided sign test excluding tied pairs.
A deterministic bootstrap 95\% confidence interval for the median difference is
computed with seed 20260621 and 10,000 resamples.

\textbf{RQ3 (CC vs.\ Test Quality).}
We compute one-sided Spearman rank correlations (left-tailed) between cyclomatic
complexity and each of branch coverage and mutation score, separately for Java and
Python.
The four planned pairs yield three testable correlations after excluding Python
mutation (which is \texttt{NOT\_TESTABLE} due to the pytest-mutagen tool
limitation).
The three raw $p$-values are adjusted using Holm--Bonferroni correction.
RQ3 is pre-registered as a secondary exploratory analysis; we do not claim causal
relationships.

% ─────────────────────────────────────────────────────────────────────────────
\subsection{Protocol Amendments and Replication Package}
\label{subsec:amendments}
% Written by Nghi (PL) — Project Lead
% ─────────────────────────────────────────────────────────────────────────────

\textbf{Protocol amendments.}
Three deviations from the registered protocol were recorded in \texttt{notes.md}
before any metric outcome was inspected.
First, an initial pilot conducted on 2026-07-02 used the GitHub Models proxy
(response IDs prefixed \texttt{mock-github-*}); this run was classified as a dry
run and excluded from analysis.
The full confirmatory run used the OpenAI API directly, producing authentic
\texttt{resp\_...} response IDs.
Second, the target model was changed from \texttt{gpt-4o-2024-08-06} to
\texttt{gpt-4o-mini-2024-07-18} for two reasons: (a)~to reduce API cost to a
total of USD\,0.065 for 50 units, and (b)~to increase output determinism by
setting \texttt{temperature}\,=\,0 (reduced from 0.2 in the proposal).
Third, on 2026-07-16, the prompt was updated to supply the full production-file
context---up to 200 lines of source---as the \texttt{MINIMAL\_CONTEXT} field,
resolving the class-name and import failures observed in the 2026-07-13 run.
Only the 2026-07-16 outputs are used in the confirmatory analysis.
All amendments and their rationales were recorded prior to the metric computation
pipeline being executed.

\textbf{Replication package.}
Our replication package includes: the frozen dataset manifest
(\texttt{experiment/dataset-manifest.csv}) with SHA-256 source hashes and
commit SHAs for all 50 units; prompt templates and their SHA-256 fingerprints;
raw OpenAI request and response JSON for all 50 API calls; dependency lock files
(\texttt{requirements-experiment.txt}) and Java build configuration; the
pre-registered analysis script; raw JaCoCo, PIT, coverage.py, and Randoop output
logs; and the full amendment log.
The package is publicly available at:
\url{https://github.com/phamvohungnghi/SWT301-SE1944-Group6}.
