% Written by Nghi (PL) — Project Lead
\section{Conclusion}
\label{sec:conclusion}

Both Java and Python cleared the pre-registered threshold with room to spare. On the Java side, 23 out of 24 executable units finished above 75\% branch coverage. The sign test returned $p = 0.000001$ after Holm adjustment, so the null hypothesis is rejected with no ambiguity. Python showed a median coverage of 100\%, with 19 of 22 executable units crossing the threshold.

The head-to-head with Randoop was the most decisive result of the study. Across all 24 matched Java pairs, the GPT-generated tests killed far more mutants than Randoop did: median mutation score of 86.6\% for GPT versus 14.8\% for Randoop (median difference $+62.1$ percentage points, Wilcoxon $p = 6.6 \times 10^{-5}$, Cliff's $\delta = 0.79$). Not a single unit went the other way. Randoop struggled on functions requiring domain-specific valid inputs (such as \texttt{isValidEmail}, \texttt{evaluateRPN}, and \texttt{jsonPathLookup}), whereas GPT's semantic understanding allowed it to construct meaningful inputs effortlessly.

Our results are broadly consistent with prior literature but benefit from greater control. Unlike TestPilot~\cite{schafer2024testpilot} or AgoneTest~\cite{lops2025agonetest}, which evaluated arbitrary code without filtering complexity or allowed multi-turn repair, our study shows that supplying full production-file context upfront in a single zero-shot prompt achieves superior first-attempt quality within the CC~5 to 15 range.

For development teams, a single call to \texttt{gpt-4o-mini-2024-07-18} with full source context produces tests hitting above 95\% branch coverage at a cost of under \$0.08 for 50 functions. Tool builders should note that the primary failure mode (string truncation) can be eliminated by implementing an automatic retry at 4,096 max tokens when syntax errors are detected.

Our complete replication package, including dataset manifests, raw API request/response logs, and pre-registered analysis scripts, is publicly available at \url{https://github.com/phamvohungnghi/SWT301-SE1944-Group6}.
